---
title: "Wrangle Democratic Candidate data"
author: "John Doe"
output: github_document
---

```{r setup, include=FALSE}
# create image folder ----
if (!file.exists("figs/")) {
    dir.create("figs/")
}
# create data folder ----
if (!file.exists("data/")) {
    dir.create("data/")
}
knitr::opts_chunk$set(
    echo = TRUE, # show all code
    eval = TRUE,
    message = FALSE, 
    comment = "#> ",
    warning = TRUE,
    tidy = FALSE, # cleaner code printing
    size = "small",
    fig.path = "figs/") # smaller code
knitr::opts_knit$set(
    width = 78)
base::options(tibble.print_max = 25,
              tibble.width = 78)
```

# Motivation 

This document describes the steps used to wrangle the candidate data.


## Packages

The Google search trends are accessible via the [gtrendsR](https://github.com/PMassicotte/gtrendsR) package. This and other packages are in the `00-packages.R` script. We load these packages in the code chunk below. 

```{r 00-packages.R, eval=TRUE}
fs::dir_tree(".")
library(gtrendsR)
library(maps)
library(ggplot2)
library(lettercase)
library(viridis)
library(pals)
library(scico)
library(ggrepel)
library(tidyverse)
library(skimr)
```

## Import data

```{r import-data}
source("src/01-import.R")
ls()
```



### Some R programming

Your code will always be communicating to at least two audiences: your computer, and your future self. Be nice to both of them!

Things like the pipe `%>%` in R can help with clarity. It takes code written like this:

```r
outer_function(inner_function(Data_X), Data_Y)
```

And makes it look like this:

```r
Data_X %>% # do this 
   inner_function() %>% # then do this
   outer_function(Data_Y)
```

## Wrangle the Wikipedia table

This table needed new names, two rows removed from the top of the table, and a numeric `airtime` variable. 

```{r WikiDemAirTime01}
# wrangle wikipedia data --------
WikiDemAirTime01 <- WikiDemAirTime01Raw %>% 
  magrittr::set_colnames(value = c("candidate", "airtime")) %>% 
  dplyr::filter(candidate %nin% c("Night one airtime", "Candidate")) %>% 
  dplyr::mutate(airtime = as.numeric(airtime))
```

## Wrangle the Google data 

The exported data is a `list`, so the `inerest_over_time` table is converted to a tibble for easier manipulation.

We then make `hits` numeric, and join the two tibbles together. 

```{r}
# convert Dems2020Group1 to tibble
Dems2020Group1IOT <- Dems2020Night1Group1$interest_over_time %>% as_tibble()
# convert Dems2020Group2 to tibble
Dems2020Group2IOT <- Dems2020Night1Group2$interest_over_time %>% as_tibble()

# create numeric hits 
Dems2020Group1IOT <- Dems2020Group1IOT %>% 
  dplyr::mutate(hits = as.numeric(hits)) 
Dems2020Group2IOT <- Dems2020Group2IOT %>%
  dplyr::mutate(hits = as.numeric(hits)) 

# bind -----------------------------------------------
Dems2020Debate01IOT <- bind_rows(Dems2020Group1IOT, 
          Dems2020Group2IOT,
          .id = "data") 
```

### Create gender

This creates the `gender` variable. 

```{r gender}
Dems2020Debate01IOT <- Dems2020Debate01IOT %>% 
  dplyr::mutate(gender = case_when(
    stringr::str_detect(keyword, "Elizabeth Warren") ~ "Women", 
    stringr::str_detect(keyword, "Amy Klobuchar") ~ "Women",
    stringr::str_detect(keyword, "Tulsi Gabbard") ~ "Women",
    TRUE ~ "Men"))
```

### Distinct rows

This will remove any duplicate rows

```{r distinct}
Dems2020Debate01IOT <- Dems2020Debate01IOT %>% distinct()
```

### Join data sources

Now we arrange the Wikipedia data by candidate and add an id column.

```{r arrange-by-candidate}
WikiDemAirTime01 <- WikiDemAirTime01 %>% dplyr::arrange(desc(candidate))
# add id
WikiDemAirTime01 <- WikiDemAirTime01 %>% 
  mutate(candidate_id = row_number())
```

Then we create an `candidate_id` column in the interest over time table. 

```{r candidate_id-IOT}
Dems2020Debate01IOT <- Dems2020Debate01IOT %>% 
  dplyr::mutate(candidate_id = case_when(
    stringr::str_detect(string = keyword, pattern = "Warren") ~ 1,
    stringr::str_detect(string = keyword, pattern = "Ryan") ~ 2,
    stringr::str_detect(string = keyword, pattern = "Beto") ~ 3,
    stringr::str_detect(string = keyword, pattern = "Klobuchar") ~ 4,
    stringr::str_detect(string = keyword, pattern = "Inslee") ~ 5,
    stringr::str_detect(string = keyword, pattern = "Gabbard") ~ 6,
    stringr::str_detect(string = keyword, pattern = "Delaney") ~ 7,
    stringr::str_detect(string = keyword, pattern = "de Blasio") ~ 8,
    stringr::str_detect(string = keyword, pattern = "Castro") ~ 9,
    stringr::str_detect(string = keyword, pattern = "Booker") ~ 10)) %>% 
  dplyr::arrange(desc(candidate_id))
```

Finally was can join.

```{r Dems2020Debate01IOTAirTime}
Dems2020Debate01IOTAirTime <- Dems2020Debate01IOT %>% 
  dplyr::left_join(x = ., 
                   y = WikiDemAirTime01, 
                   by = "candidate_id")
```


### Prior approval percent of voters

This is the prior percent approval from the [MorningConsult](https://morningconsult.com/) survey on [fivethirtyeight](https://projects.fivethirtyeight.com/democratic-debate-poll/), but these data are no longer avaiable. 

> To track which candidates are winning over voters, we asked respondents who they would vote for before and after each debate. That lets us measure not only who gained (or lost) support, but also where that support came from (or went to).

```{r prior_vperc}
# prior_vperc -------------------------------------------------------------
Dems2020Debate01IOTAirTime <- Dems2020Debate01IOTAirTime %>% 
  dplyr::mutate(prior_vperc = case_when(
    stringr::str_detect(keyword, "O’Rourke") ~ "> 1.0% of voters",
    stringr::str_detect(keyword, "Warren") ~ "> 1.0% of voters",
    stringr::str_detect(keyword, "Booker") ~ "> 1.0% of voters",
    
    stringr::str_detect(keyword, "Klobuchar") ~ "0.5 - 0.9% of voters",
    stringr::str_detect(keyword, "Castro") ~ "0.5 - 0.9% of voters",

    stringr::str_detect(keyword, "Gabbard") ~ "0.2 - 0.4% of voters",
    stringr::str_detect(keyword, "Ryan") ~ "0.2 - 0.4% of voters",
    stringr::str_detect(keyword, "Inslee") ~ "0.2 - 0.4% of voters",
    stringr::str_detect(keyword, "de Blasio") ~ "0.2 - 0.4% of voters",

    stringr::str_detect(keyword, "Delaney") ~ "0.2% of voters"))
# display tabyl
Dems2020Debate01IOTAirTime %>% 
  dplyr::count(prior_vperc, keyword) %>% 
  tidyr::spread(prior_vperc, n)
```

Then turn this into a factor and check the levels. 

```{r prior_vperc_fct}
# assing labels
Dems2020Debate01IOT <- Dems2020Debate01IOT %>% 
  dplyr::mutate(prior_vperc_fct = factor(x = prior_vperc))

Dems2020Debate01IOT <- Dems2020Debate01IOT %>% 
  dplyr::mutate(prior_vperc_fct = forcats::fct_relevel(.f = prior_vperc_fct,
                    "> 1.0% of voters",
                    "0.5 - 0.9% of voters",
                    "0.2 - 0.4% of voters",
                    "0.2% of voters"))
# check
Dems2020Debate01IOT$prior_vperc_fct %>% levels()
```

## Create the map data

We have to go back to the original exported list `Dems2020Night1Group1` to get the mapping data put together.  

```{r Dems2020IBR}
# convert to tibble (another data structure in R)
Dems2020IBRGroup1 <- tibble::as_tibble(Dems2020Night1Group1$interest_by_region)
Dems2020IBRGroup2 <- tibble::as_tibble(Dems2020Night1Group2$interest_by_region)
# bind Dems2020IBRGroup1 Dems2020IBRGroup2 together 
Dems2020IBR <- bind_rows(Dems2020IBRGroup1, 
                              Dems2020IBRGroup2, .id = "data")
```

Next we convert the `region` variable to lowercase and join this to the `statesMap` data from `ggplot2` 

```{r Dems2020IBRMapData}
# convert the region to lowercase
Dems2020IBR <- Dems2020IBR %>% 
  dplyr::mutate(region = stringr::str_to_lower(location))
# create a data set for the states in the US
statesMap = ggplot2::map_data("state")
# now merge the two data sources together
Dems2020IBRMapData <- Dems2020IBR %>% 
  dplyr::inner_join(x = .,
                   y = statesMap, 
                   by = "region")
```

Now we have two data sets for visualizations and maps!
